-
title: "CISC 351 Project"
output: rmarkdown::github_document
---

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

Read in datasets
```{r}
library(readxl)
adv_stats <- read_excel("C:/Users/Quentin/Desktop/Dataset project/NBA Players - Advanced Season Stats (1978-2016).xlsx")
basic_stats <- read_excel("C:/Users/Quentin/Desktop/Dataset project/NBA Players - Basic Season Stats (1950-2017).xlsx")
salary_cap <- read_excel("C:/Users/Quentin/Desktop/Dataset project/nba_salary_cap.xlsx", col_names = TRUE)

```

Data cleaning - Compile datasets
```{r}
# Remove empty/unnecessary columns
basic_sub = as.data.frame(basic_stats[, -c(1, 23, 28)])
adv_sub = as.data.frame(adv_stats[, -c(2, 9, 19, 24, 32, 35)])
# Rename columns in basic_sub for arrange() and filter()
colnames(basic_sub)[1] = 'Year'
colnames(basic_sub)[2] = 'Player'
colnames(basic_sub)[3] = 'Salary'
# Create consistent ordering for both datasets
library(tidyverse)
basic = filter(arrange(basic_sub, Year, Player), Year < 2016 & Year > 1989)
adv = filter(arrange(adv_sub, year, player), year < 2016 & year > 1989)
# Remove observations that split a player's season stats before and after a trade has occurred
index = rep(TRUE, nrow(basic))
for (i in 1:(nrow(basic) - 1)) {
  if (basic$Year[i] == basic$Year[i + 1] &  basic$Player[i] == basic$Player[i + 1]) {
    index[i + 1] = FALSE
  }
}
basic2 = basic[index,]
index2 = rep(TRUE, nrow(adv))
for (i in 1:(nrow(adv) - 1)) {
  if (adv$year[i] == adv$year[i + 1] &  adv$player[i] == adv$player[i + 1]) {
    index2[i + 1] = FALSE
  }
}
adv2 = adv[index2,]
# Final observations with matching names
basic2[-c(6764, 8726, 8730, 8745, 9644, 10028), ]
adv2[-c(8604, 8786, 9046), ]
```

Data Cleaning - Remove recurrent columns and observations without recorded salary
```{r}
# Prune redundant covariates from basic data
basic = basic2[-c(6764, 8726, 8730, 8745, 9644, 10028), c(1:5, 7:29, 32, 35, 38, 39, 42:44, 46:51)]
# Prune redundant covariates from advanced data
advanced = adv2[-c(8604, 8786, 9046), c(23, 31, 33, 35:37, 39, 85, 86, 88)]
data = cbind(basic, advanced)  ## Join data sets
data = data[!is.na(data$Salary),]  ## Remove observations without salary
```
Feature Engineering - Adjust points, rebounds, assists, blocks to per game metric and salary to percentage
```{r}
# Convert salary feature to percentage of total salary cap and standardize remaining features
for (i in 1:nrow(data)) {
  data$Salary[i] = data$Salary[i] / salary_cap[salary_cap$Year == data$Year[i], 2]
  data$PTS[i] = data$PTS[i] / data$G[i]
  data$AST[i] = data$AST[i] / data$G[i]
  data$STL[i] = data$STL[i] / data$G[i]
  data$DRB[i] = data$DRB[i] / data$G[i]
  data$ORB[i] = data$ORB[i] / data$G[i]
  data$BLK[i] = data$BLK[i] / data$G[i]
  data$TOV[i] = data$TOV[i] / data$G[i]
  data$PF[i] = data$PF[i] / data$G[i]
  data$MP[i] = data$MP[i] / data$G[i]
  if (is.na(data$`3P%`[i])) {
    data$`3P%`[i] = 0
  }
  if (is.na(data$`FT%`[i])) {
    data$`FT%`[i] = 0
  }
}
# Position as factor
data$Pos = as.factor(data$Pos)
# Remove observations with missing values
na = rep(0, ncol(data))
for (i in 1:ncol(data)) {
  na[i] = sum(is.na(data[, i]))
}
#Renaming certain factors for ease with randomForest library
colnames(data)[10] = 'TS.percentage'
colnames(data)[11] = 'threePAr'
colnames(data)[13] = 'ORB.percentage'
colnames(data)[14] = 'DRB.percentage'
colnames(data)[15] = 'TRB.percentage'
colnames(data)[16] = 'AST.percentage'
colnames(data)[17] = 'STL.percentage'
colnames(data)[18] = 'BLK.percentage'
colnames(data)[19] = 'TOV.percentage'
colnames(data)[20] = 'USG.percentage'
colnames(data)[24] = 'WS.on.forty.eight'
colnames(data)[29] = 'FG.percentage'
colnames(data)[30] = 'three.point.percentage'
colnames(data)[31] = 'two.point.percentage'
colnames(data)[32] = 'eFG.percentage'
colnames(data)[33] = 'FT.percentage'

data$Salary = as.numeric(data$Salary)
```
Random Forest setup
```{r}
library(randomForest)
set.seed(1)
train <- sample(nrow(data), 0.8*nrow(data), replace = FALSE)
trainSet <- data[train,]
testSet <- data[-train,]

model1 <- randomForest(Salary ~ ., data = trainSet, importance = TRUE, na.action = na.omit)
model1
plot (model1)
which.min(model1$mse)
sqrt(model1$mse[which.min(model1$mse)])

model2 <- randomForest(Salary ~ ., data = trainSet, ntree = 424, importance = TRUE, na.action = na.omit)
which.min(model2$mse)
sqrt(model2$mse[which.min(model2$mse)])
model2

for(i in 1:15) {
  model3 <- randomForest(Salary ~ ., data = trainSet, ntree = 424, mtry = (5 + i),  importance = TRUE, na.action = na.omit)
  print(i+5)
  model3
  print(sqrt(model3$mse[which.min(model3$mse)]))
}
```